Short-term neuronal and synaptic plasticity act in synergy for deviance detection in spiking networks

### 2023-03-20 paper-1/elife.tex
Better fig name: 2 -> DD

\begin{figure}
    \includegraphics[width=\linewidth]{fig2}
    \caption{%
        Deviance detection in our model.
        \textbf{A} Sample network target response intensity by trial type. Each horizontal line of the plot summarizes the network response in one target trial in terms of spikes per integration time step. Target trials are taken from the \std{} (400), \msc{} (100), and \dev{} (100) sequences, and sorted by sequence time (earliest trials are at the bottom of each section).
        \textbf{B} Response summary, showing average network-wide spike counts in target trials in each sequence. The underlying data are the same as in panel A. Note how the \dev{} response is very large, followed by an intermediate \msc{} response and a low \std{} response.
        \textbf{C} Mean response sizes $\mean{\R[dev]{A}{}}$ and $\mean{\R[std]{A}{}}$, normalized by $\mean{\R[msc]{A}{}}$, across networks and stimuli (n = 120). In this and later boxplots, the median is indicated with an orange line, notches indicate the 95\% confidence interval for the median calculated by bootstrap with 10000 iterations, boxes indicate the inter-quartile range (IQR), whiskers extend to 1.5 times the IQR or to the data extrema, whichever is less, and fliers indicate individual data points beyond the whisker end points.
        \textbf{D} Deviance detection indices (DDI, see Equation {eq:ddi}) across networks and stimuli.
    }
    \label{fig:2}
\end{figure}

To show that our model is capable of deviance detection, we first show the responses of the sample network in all sequences (\dev{}, \msc{}, and \std{}) in \FIG{2}A. Here, each row corresponds to one trial, aligned to stimulus input at t = 0, with the number of spikes in each millisecond bin indicated by color. All responses share an early activity pattern, which derives from the stimulated neurons and their immediate postsynaptic targets, but differences soon start to emerge. The \dev{} sequence responses, shown at the top of the plot, exhibit a clear peak in activity around 15 ms after stimulation. The \std{} sequence responses show a later, more diffuse activity peak around 20 ms, and the \msc{} sequence responses lack any obvious activity cluster. We sum these patterns up in \FIG{2}B, where we show the mean network activity, in terms of spikes per millisecond bin, across trials of each type. Clearly, the target response in the \dev{} sequence is greater than either the \msc{} or the \std{} response.

We could readily identify similar patterns emerging across most networks and stimulus assignments. To show that deviance detection is robust to network and stimulus identity, we calculated the average spike count per neuron per target trial for each network and stimulus assignment, and compared these across sequences. \FIG{2}C shows a summary of the average \dev{} and \std{} response sizes, normalized by the response in the \msc{} sequence. The data clearly confirmed that \dev{} responses were larger than \msc{} (T = 6.38e+03, p = 2.75e-13; one-sided Wilcoxon signed-rank test, n = 120), and \std{} responses smaller (T = 687, p = 6.42e-15). Accordingly, the DDI (\FIG{2}D) was positive in most networks (T = 5.97e+03, p = 4.75e-10, median 0.114). This shows that our model was well behaved and suitable for an investigation of deviance detection.

>>>>

\begin{figure}
    \includegraphics[width=\linewidth]{fig-dd}
    \caption{%
        Deviance detection in our model.
        \textbf{A} Sample network target response intensity by trial type. Each horizontal line of the plot summarizes the network response in one target trial in terms of spikes per integration time step. Target trials are taken from the \std{} (400), \msc{} (100), and \dev{} (100) sequences, and sorted by sequence time (earliest trials are at the bottom of each section).
        \textbf{B} Response summary, showing average network-wide spike counts in target trials in each sequence. The underlying data are the same as in panel A. Note how the \dev{} response is very large, followed by an intermediate \msc{} response and a low \std{} response.
        \textbf{C} Mean response sizes $\mean{\R[dev]{A}{}}$ and $\mean{\R[std]{A}{}}$, normalized by $\mean{\R[msc]{A}{}}$, across networks and stimuli (n = 120). In this and later boxplots, the median is indicated with an orange line, notches indicate the 95\% confidence interval for the median calculated by bootstrap with 10000 iterations, boxes indicate the inter-quartile range (IQR), whiskers extend to 1.5 times the IQR or to the data extrema, whichever is less, and fliers indicate individual data points beyond the whisker end points.
        \textbf{D} Deviance detection indices (DDI, see Equation {eq:ddi}) across networks and stimuli.
    }
    \label{fig:DD}
\end{figure}

To show that our model is capable of deviance detection, we first show the responses of the sample network in all sequences (\dev{}, \msc{}, and \std{}) in \FIG{DD}A. Here, each row corresponds to one trial, aligned to stimulus input at t = 0, with the number of spikes in each millisecond bin indicated by color. All responses share an early activity pattern, which derives from the stimulated neurons and their immediate postsynaptic targets, but differences soon start to emerge. The \dev{} sequence responses, shown at the top of the plot, exhibit a clear peak in activity around 15 ms after stimulation. The \std{} sequence responses show a later, more diffuse activity peak around 20 ms, and the \msc{} sequence responses lack any obvious activity cluster. We sum these patterns up in \FIG{DD}B, where we show the mean network activity, in terms of spikes per millisecond bin, across trials of each type. Clearly, the target response in the \dev{} sequence is greater than either the \msc{} or the \std{} response.

We could readily identify similar patterns emerging across most networks and stimulus assignments. To show that deviance detection is robust to network and stimulus identity, we calculated the average spike count per neuron per target trial for each network and stimulus assignment, and compared these across sequences. \FIG{DD}C shows a summary of the average \dev{} and \std{} response sizes, normalized by the response in the \msc{} sequence. The data clearly confirmed that \dev{} responses were larger than \msc{} (T = 6.38e+03, p = 2.75e-13; one-sided Wilcoxon signed-rank test, n = 120), and \std{} responses smaller (T = 687, p = 6.42e-15). Accordingly, the DDI (\FIG{DD}D) was positive in most networks (T = 5.97e+03, p = 4.75e-10, median 0.114). This shows that our model was well behaved and suitable for an investigation of deviance detection.

### 2023-03-20 paper-1/manuscript.tex
Better fig name: 2 -> DD

\begin{figure*}%
    \centering
    \includegraphics{fig2.pdf}
    \caption{}
    \label{fig:2}
\end{figure*}
\textbf{Fig. 2} Deviance detection in our model.
\textbf{a} Sample network target response intensity by trial type. Each horizontal line of the plot summarizes the network response in one target trial in terms of spikes per integration time step. Target trials are taken from the \std{} (400), \msc{} (100), and \dev{} (100) sequences, and sorted by sequence time (earliest trials are at the bottom of each section).
\textbf{b} Response summary, showing average network-wide spike counts in target trials in each sequence. The underlying data are the same as in panel a. Note how the \dev{} response is very large, followed by an intermediate \msc{} response and a low \std{} response.
\textbf{c} Mean response sizes $\mean{\R[dev]{A}{}}$ and $\mean{\R[std]{A}{}}$, normalized by $\mean{\R[msc]{A}{}}$, across networks and stimuli (n = 120). In this and later boxplots, the median is indicated with an orange line, notches indicate the 95\% confidence interval for the median calculated by bootstrap with 10000 iterations, boxes indicate the inter-quartile range (IQR), whiskers extend to 1.5 times the IQR or to the data extrema, whichever is less, and fliers indicate individual data points beyond the whisker end points.
\textbf{d} Deviance detection indices (DDI, see Equation {eq:ddi}) across networks and stimuli

To show that our model is capable of deviance detection, we first show the responses of the sample network in all sequences (\dev{}, \msc{}, and \std{}) in \FIG{2}a. Here, each row corresponds to one trial, aligned to stimulus input at t = 0, with the number of spikes in each millisecond bin indicated by color. All responses share an early activity pattern, which derives from the stimulated neurons and their immediate postsynaptic targets, but differences soon start to emerge. The \dev{} sequence responses, shown at the top of the plot, exhibit a clear peak in activity around 15 ms after stimulation. The \std{} sequence responses show a later, more diffuse activity peak around 20 ms, and the \msc{} sequence responses lack any obvious activity cluster. We sum these patterns up in \FIG{2}b, where we show the mean network activity, in terms of spikes per millisecond bin, across trials of each type. Clearly, the target response in the \dev{} sequence is greater than either the \msc{} or the \std{} response.

We could readily identify similar patterns emerging across most networks and stimulus assignments. To show that deviance detection is robust to network and stimulus identity, we calculated the average spike count per neuron per target trial for each network and stimulus assignment, and compared these across sequences. \FIG{2}c shows a summary of the average \dev{} and \std{} response sizes, normalized by the response in the \msc{} sequence. The data clearly confirmed that \dev{} responses were larger than \msc{} (T = 6.38e+03, p = 2.75e-13; one-sided Wilcoxon signed-rank test, n = 120), and \std{} responses smaller (T = 687, p = 6.42e-15). Accordingly, the DDI (\FIG{2}d) was positive in most networks (T = 5.97e+03, p = 4.75e-10, median 0.114). This shows that our model was well behaved and suitable for an investigation of deviance detection.

>>>>

\begin{figure*}%
    \centering
    \includegraphics{fig-dd.pdf}
    \caption{}
    \label{fig:DD}
\end{figure*}
\textbf{Fig. 2} Deviance detection in our model.
\textbf{a} Sample network target response intensity by trial type. Each horizontal line of the plot summarizes the network response in one target trial in terms of spikes per integration time step. Target trials are taken from the \std{} (400), \msc{} (100), and \dev{} (100) sequences, and sorted by sequence time (earliest trials are at the bottom of each section).
\textbf{b} Response summary, showing average network-wide spike counts in target trials in each sequence. The underlying data are the same as in panel a. Note how the \dev{} response is very large, followed by an intermediate \msc{} response and a low \std{} response.
\textbf{c} Mean response sizes $\mean{\R[dev]{A}{}}$ and $\mean{\R[std]{A}{}}$, normalized by $\mean{\R[msc]{A}{}}$, across networks and stimuli (n = 120). In this and later boxplots, the median is indicated with an orange line, notches indicate the 95\% confidence interval for the median calculated by bootstrap with 10000 iterations, boxes indicate the inter-quartile range (IQR), whiskers extend to 1.5 times the IQR or to the data extrema, whichever is less, and fliers indicate individual data points beyond the whisker end points.
\textbf{d} Deviance detection indices (DDI, see Equation {eq:ddi}) across networks and stimuli

To show that our model is capable of deviance detection, we first show the responses of the sample network in all sequences (\dev{}, \msc{}, and \std{}) in \FIG{DD}a. Here, each row corresponds to one trial, aligned to stimulus input at t = 0, with the number of spikes in each millisecond bin indicated by color. All responses share an early activity pattern, which derives from the stimulated neurons and their immediate postsynaptic targets, but differences soon start to emerge. The \dev{} sequence responses, shown at the top of the plot, exhibit a clear peak in activity around 15 ms after stimulation. The \std{} sequence responses show a later, more diffuse activity peak around 20 ms, and the \msc{} sequence responses lack any obvious activity cluster. We sum these patterns up in \FIG{DD}b, where we show the mean network activity, in terms of spikes per millisecond bin, across trials of each type. Clearly, the target response in the \dev{} sequence is greater than either the \msc{} or the \std{} response.

We could readily identify similar patterns emerging across most networks and stimulus assignments. To show that deviance detection is robust to network and stimulus identity, we calculated the average spike count per neuron per target trial for each network and stimulus assignment, and compared these across sequences. \FIG{DD}c shows a summary of the average \dev{} and \std{} response sizes, normalized by the response in the \msc{} sequence. The data clearly confirmed that \dev{} responses were larger than \msc{} (T = 6.38e+03, p = 2.75e-13; one-sided Wilcoxon signed-rank test, n = 120), and \std{} responses smaller (T = 687, p = 6.42e-15). Accordingly, the DDI (\FIG{DD}d) was positive in most networks (T = 5.97e+03, p = 4.75e-10, median 0.114). This shows that our model was well behaved and suitable for an investigation of deviance detection.
